---
title: "Major_1_Tingyu_Liu1003"
author: "Drunken_Boat"
date: "2023-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(here)
library(tidycensus)
library(readr)
library(skimr)
library(ggplot2)
library(lubridate)
library(dplyr)
library(htmlwidgets)
```

```{r 1_load_data, message=FALSE, warning=FALSE}
data <- read_csv("fe.csv")
head(data)
```

```{r 2_eda}
skim(data)
```

```{r 3_gender_distribution}
p_distribution <- ggplot(data = data) +
  geom_bar(aes(x = "", fill = Gender), stat = "count", position = "fill") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = c("#00008B","#AA0000", "#333333","#999999"))+
  theme(panel.background = element_rect(fill = "black"))+
  ggdark::dark_theme_classic()

print(p_distribution)
```
- The pie chart above illustrates the distribution of police killings based on gender.
- The majority of police killings were of Male, accounting for nearly ninety percent of the total incidents.

```{r 4_age_and_gender}
p_age <- ggplot(data) +
  geom_histogram(mapping = aes(x = Age, fill = Gender), 
                 bins = 60,
                 color="black")+
  scale_fill_manual(values = c("#00008B","#AA0000", "#333333","#999999"))+
  ggdark::dark_theme_classic()
print(p_age)
```
- The stacked bar chart depicts the distribution of police killings based on both gender and age groups.
- The highest number of incidents occurred in the 25-40 group.
- There is a noticeable disparity in police violence across different age ranges and genders, with young adult female.

```{r 5_trend_over_time}
colnames(data)[9] <- "time"

data_time <- data %>%
  mutate(time = as.Date(time, format = "%m/%d/%Y"))

data_time <- data_time %>%
  mutate(year = year(time))

# Count the number of incidents per year and gender
incident_counts <- data_time %>%
  group_by(year, Gender) %>%
  count()

color_palette <- c("Female" = "#00008B", "Male" = "#AA0000", "Transgender" = "#333333", "NA" = "#00000000")

p_trend <- ggplot(data = incident_counts) +
  geom_line(aes(x = year, y = n, color = Gender), size=2)+
  scale_color_manual(values = color_palette) +
  labs(x = "Year", y = "Number of Incidents")+ 
  ggdark::dark_theme_classic()

print(p_trend)
```
- The line chart above showcases the trend of police killings over time.
- There has been an obvious increase trend in police killings over the years, from 2000 - 2021.
- The number of incidents peaked in 2020.

```{r 6_race}
data_race <- read_csv("fe-cleaned.csv")

p_race <- ggplot(data = data_race) + 
  geom_bar(mapping = aes(x=Gender, fill=Race), position = "fill") +
  scale_fill_manual(values = c("#AA0000", "#FF8C00", "#006400", "#000080", "#008B8B", "#006400", "#333333"))+
  ggdark::dark_theme_classic()

print(p_race)
```
- The stacked bar chart above shows the distribution of police killings based on both gender and race groups.
- Among the Female group, more victims Asian and White victims occurs compared to other genders.
```{r 7_save}
save.image('1005.RData')
```



